---
title: "NYPD_project_pacheco_eli"
output:
  pdf_document: default
  html_document: default
date: "2023-10-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ADDED STEP:  Import Libraries
```{r library, message=FALSE, warning=FALSE}

library(tidyverse)
library(lubridate)
```


## STEP 1: Start an Rmd document that describes and imports the shooting project dataset in a reproducible manner.

Using a simple url_in method i will import data and assign a variable that is reproducible.

```{r import_data, message=FALSE}
url_in = "https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD"
data_set = read_csv(url_in)
```

## STEP 2: Tidy and Transform Your Data
There are a number of columns i will not be needing for my assessment, thus, i will remove them and tidy my data.

```{r}
data_set2 = data_set %>% 
  select(-c(INCIDENT_KEY,PRECINCT, JURISDICTION_CODE, LOC_CLASSFCTN_DESC, LOC_OF_OCCUR_DESC, LOCATION_DESC, X_COORD_CD, Y_COORD_CD, Longitude, Latitude, Lon_Lat))
print(data_set2)
summary(data_set2)
```
In this figure, the columns not needed have been taken out of the final data set.


## STEP 3: Add Visualizations and Analysis
The first question I have is:  *Which neighborhoods are the most dangerous over the course of time.*
I will make a basic bar graph to illustrate this.

The second question is: *At what time of the day these tend to occur.*
I will make a frequency chart to display that data.

The final question I have is:  *Which days are most likely to have an occurence.*
This will be displayed in a basic bar graph.

These three inquiries will be useful for anyone wanting to visit these neighborhoods. 

```{r}
graph1 = ggplot(data_set2, aes(x = BORO)) +
  geom_bar() +
  labs(title = "Boroughs of New York City with shooting stats",
       x = "Boroughs of New York City",
       y = "Count of Incidents") +
  theme_minimal()
graph1

ggplot(data = data_set2, aes(x = OCCUR_TIME)) +
  geom_freqpoly(bins = 50)+
  labs(title = "Peak Dangerous Times",
       x = "Time of Occurrence")

data_set2$OCCUR_DAY = mdy(data_set2$OCCUR_DATE)
data_set2$OCCUR_DAY = wday(data_set2$OCCUR_DAY, label = TRUE)
data_set2$OCCUR_HOUR = hour(hms(as.character(data_set2$OCCUR_TIME)))

x1 = data_set2 %>%
  group_by(OCCUR_DAY) %>%
  count()

g <- ggplot(x1, aes(x = OCCUR_DAY, y = n)) +
  geom_col() +
  labs(title = "Most likely days of Occurrences",
       x = "Incident Occurence Day",
       y = "Count of Incidents") +
  theme_minimal()
g

```
```{r}
g <- ggplot(df_4, aes(x = OCCUR_HOUR, y = n)) +
  geom_line() +
  labs(title = "Which time should people in New York be cautious of incidents?",
       x = "Incident Occurence Hour",
       y = "Count of Incidents") +
  theme_minimal()
```

*CONCLUSION FROM STEP 3*
Conclusion 1:  Brooklyn and the Bronx are most likely areas of shootings, with Staten Island being least likely.
Conclusion 2:  Time of occurrences spike during the evening/nights.
Conclusion 3:  Occurrences spike during this weekends. 

This information can be very helpful for anyone looking to visit these neighborhoods.  This can be used in a number of modern tourism applications. 
Ensuring some level of safety is important when touring any new city/neighborhood. 

## Step 4 Add Bias Identification

Data used in this report was collected from the City of New York.  Collection methods by the Police Department may be an inherit bias.  A helpful statistic would be knowing the amount of police force in each neighborhood during these times to correlate any ties to surge police activity and police reports.  
It must also be noted that I have visited these neighborhoods many times in my life, and have formulated opinions on each neighborhood already.  In order to mitigate this, I created a simple bar graph to show each neighborhood statistically, rather than push my own bias further. 




